---
title: "DataFormat Code"
author: "Saulai Vue"
date: "2025-05-13"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
library(stringr)
library(conflicted)
library(dplyr)
library(tidyverse)
library(tidyr)
library(rvest) #web scraping package and working with html
library(stringr) #working with strings into tidyverse and cleaning text strings
library(readr)
library(httr)
```

```{r}
scraped_lyrics_url <- "https://raw.githubusercontent.com/vsaulai/STAT-345-FinalProject-Lyrics/refs/heads/main/webscraping/Lyrics_updated_repeatcount.csv"
scraped_lyrics_df <- read.csv(scraped_lyrics_url)
View(scraped_lyrics_df)
```


## Function to Remove Punctuation, "[Instrumental]", and Invalid entries

```{r}
remove_brackets_clean_text <- function(data, column_name) {
  if (!column_name %in% names(data)) {
    stop("The specified column does not exist in the data frame.")
  }

  # Fix encoding issues
  data[[column_name]] <- iconv(data[[column_name]], to = "UTF-8", sub = "byte")

  # Replace values that start with "We do not have the lyrics" (case-insensitive) with NA
  data[[column_name]][grepl("^we do not have the lyrics", data[[column_name]], ignore.case = TRUE)] <- NA

  # Remove content in brackets (), [], {}
  data[[column_name]] <- gsub("\\[[^\\]]*\\]|\\([^\\)]*\\)|\\{[^\\}]*\\}", "", data[[column_name]])

  # Remove specific words like "Instrumental" (case-insensitive)
  data[[column_name]] <- gsub("\\bInstrumental\\b", "", data[[column_name]], ignore.case = TRUE)

  # Remove punctuation except apostrophes
  data[[column_name]] <- gsub("[^[:alnum:] [:space:]']", "", data[[column_name]])

  # Trim extra white space
  data[[column_name]] <- trimws(data[[column_name]])

  return(data)
}

```

## Count number of words column

```{r}
add_word_count_column <- function(data, column_name, new_column_name = "word_count") {
  if (!column_name %in% names(data)) {
    stop("The specified column does not exist in the data frame.")
  }

  # Handle NA safely, count words using strsplit
  word_counts <- sapply(data[[column_name]], function(x) {
    if (is.na(x) || trimws(x) == "") {
      return(0)
    } else {
      return(length(strsplit(trimws(x), "\\s+")[[1]]))
    }
  })

  # Add the new word count column
  data[[new_column_name]] <- word_counts

  return(data)
}

```

## Run Previous Functions

```{r}
scraped_lyrics_cleaning1 <- remove_brackets_clean_text(scraped_lyrics_df, "lyrics")
View(scraped_lyrics_cleaning1)
```

```{r}
cleaning_url <- "https://raw.githubusercontent.com/vsaulai/STAT-345-FinalProject-Lyrics/refs/heads/main/cleaning_lyrics.csv"
cleaning_df <- read.csv(cleaning_url)
View(cleaning_df)
```

```{r}
counting_lyrics <- add_word_count_column(cleaning_df, "lyrics")
View(counting_lyrics)
```


```{r}
lyrics_df <- counting_lyrics
```

## Installing Udpipe
```{r}
install.packages("udpipe")
library(udpipe)
```


## Udpipe Language Model

```{r}
model <- udpipe_download_model(language = "english")
ud_model <- udpipe_load_model(model$file_model)
```

## Function to Count word type

```{r}
count_pos_types <- function(df, column_name, ud_model) {
  # Store results in a list
  pos_counts_list <- lapply(df[[column_name]], function(text) {
    if (is.na(text) || trimws(text) == "") return(data.frame())
    
    # Annotate the text
    annotation <- udpipe_annotate(ud_model, x = text)
    anno_df <- as.data.frame(annotation)
    
    # Count POS tags
    table(anno_df$upos)
  })
  
  # Convert to a tidy data frame
pos_summary <- do.call(rbind, lapply(seq_along(pos_counts_list), function(i) {
  pos_counts <- pos_counts_list[[i]]
  
  if (!is.null(pos_counts) && length(pos_counts) > 0) {
    row <- as.data.frame(as.table(pos_counts))
    row$id <- i
    return(row)
  } else {
    return(data.frame(Var1 = NA, Freq = NA, id = i))
  }
}))
  # Reshape to wide format: one row per observation, POS tags as columns+
  pos_wide <- pivot_wider(pos_summary, names_from = Var1, values_from = Freq, values_fill = 0)
  
  # Join with original data frame (by row index)
  df$id_tmp <- seq_len(nrow(df))
  result <- merge(df, pos_wide, by.x = "id_tmp", by.y = "id", all.x = TRUE)
  result$id_tmp <- NULL
  
  return(result)
}
```

```{r}
result_types <- count_pos_types(lyrics_df, "lyrics", ud_model)
View(result_types)
```

## Renaming POS tags from english UDMODEL
    noun = NOUN, verb = VERB,
    adjective  = ADJ,
    adposition = ADP,
    adverb  = ADV,
    auxiliary_verb = AUX,
    proper_noun = PROPN,
    interjection = INTJ,
    numeral = NUM,
    particle = PART,
    pronoun = PRON,
    subordinating_conj = SCONJ,
    coordinating_conj = CCONJ,
    determiner = DET,
    punctuation = PUNCT,
    symbol = SYM,
    other = X,

```{r}
main_lyrics_df <- result_types %>%
  rename(
    noun = NOUN,
    verb = VERB,
    adjective  = ADJ,
    adposition = ADP,
    adverb  = ADV,
    proper_noun = PROPN,
    interjection = INTJ,
    numeral = NUM,
    particle = PART,
    pronoun = PRON,
    subordinating_conj = SCONJ,
    coordinating_conj = CCONJ,
    determiner = DET,
    punctuation = PUNCT,
    symbol = SYM,
    other = X,
  )
# Print 5 Columns 9:22 for 

head(df[9:22, ], n = 5)

```


## Code to recount title name frequency
- Function 1: Measure phrase repetition

```{r}
# Function 1: Measure phrase repetition
measure_phrase_repetition <- function(lyrics, phrase_length = 2) {
  words <- unlist(str_split(tolower(lyrics), "\\s+"))  # Split lyrics into words
  
  # Create phrase sequences
  phrases <- sapply(1:(length(words) - (phrase_length - 1)), function(i) {
    paste(words[i:(i + phrase_length - 1)], collapse = " ")
  })
  
  phrase_counts <- table(phrases)  # Count occurrences
  sorted_phrases <- sort(phrase_counts, decreasing = TRUE)  # Sort by frequency
  
  return(names(head(sorted_phrases, 10)))  # Return top 10 repeated phrases
}
```

## Function 2: Measure title repetition

```{r}
# Function 2: Measure title repetition
measure_title_repetition <- function(title, lyrics) {
  if (is.na(lyrics) || is.na(title) || str_trim(lyrics) == "") return(NA_integer_)
  
  # Try converting to lowercase safely
  title_clean <- tryCatch(tolower(title), error = function(e) return(NA_character_))
  lyrics_clean <- tryCatch(tolower(lyrics), error = function(e) return(NA_character_))
  
  if (is.na(title_clean) || is.na(lyrics_clean)) return(NA_integer_)
  
  title_count <- str_count(lyrics_clean, fixed(title_clean))
  return(title_count)
}
```

```{r}
# Load dataset
lyrics_data <- main_lyrics_df
```

```{r}
# Apply repetitiveness measures
lyrics_data <- lyrics_data %>%
  rowwise() %>%
  mutate(
    phrase_repetition = paste(measure_phrase_repetition(lyrics), collapse = "; "),
    title_repetition_count = measure_title_repetition(Title, lyrics)
  ) %>%
  ungroup()
```

```{r}
write.csv(lyrics_data, "final_lyrics_df.csv", row.names = FALSE)
```

## Main Data frame Created on Github

```{r}
lyrics_final_df <- "https://raw.githubusercontent.com/vsaulai/STAT-345-FinalProject-Lyrics/refs/heads/main/final_lyrics_df.csv"
lyrics_df <- read.csv(lyrics_final_df)
View(lyrics_df)
```

