---
title: "missingSaulaiVueCode"
author: "Saulai Vue"
date: "2025-05-07"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(stringr)
library(conflicted)
library(dplyr)
library(tidyverse)
library(tidyr)
library(rvest) #web scraping package and working with html
library(stringr) #working with strings into tidyverse and cleaning text strings
library(readr)
library(httr)
conflict_prefer("filter", "dplyr")
```

## Importing from Github MISSING 1971, 2012, 2013 missing top 20 artists and songs

```{r}
missing_df_71_12_13 <- "https://raw.githubusercontent.com/vsaulai/STAT-345-FinalProject-Lyrics/refs/heads/main/billboard_missing_1971_2012_2013.csv" %>%
                      read.csv()
View(missing_df_71_12_13)
```

## Importing MAIN Data Frame from Github (will combine to this one eventually)
- Problems with initial scraping
  - Using escape characters to define string characters for the artist and song name caused a lot of problems
  - Song collaborations caused a lot of problems when scraping on a legitimate lyrics site because the host website didn't accomadate artist features in a url, therefore we had to format out artist string names including "featurinh" "feat." "ft."
  - Duplicate Songs by the same artist were recorded in the same song entry. Had to split this entry into another row

```{r}
main_df_artist_lyrics <- "https://raw.githubusercontent.com/vsaulai/STAT-345-FinalProject-Lyrics/refs/heads/main/LYRICS_Top_20_Songs_1958_2024.csv" %>%
                        read.csv()
View(main_df_artist_lyrics)
```

## Cleaning the Missing_1971_2012_2013 artist and song list.
- Need to remove quotes from song title before generating URL

```{r}
missing_df_71_12_13$Title <- gsub("[\"']", "", missing_df_71_12_13$Title) #remove quotes from Song Title
```

## Function No.1: 'split_and_insert_row(data frame, "column_name")'
  - There are two songs in song title that belong to the same artist and are placed in the same rankings
  - This function splits this song and creates another song lyric entry

```{r}
# Function to split a column by "/" and create new rows
split_and_insert_row <- function(df, column_name) {
  # Create an empty list
  new_rows <- list()
  # Loop through each row of the dataframe
  for (i in 1:nrow(df)) {
    # Get the value in the specified column to split
    values_to_split <- df[i, column_name]
    
    # If the value contains "/", split it and create new rows
    if (grepl("/", values_to_split)) {
      split_values <- strsplit(values_to_split, "/")[[1]]
      
      # For each split value, create a new row with the same other column values
      for (split_value in split_values) {
        new_row <- df[i, , drop = FALSE]  # Keep the row structure
        new_row[[column_name]] <- split_value
        new_rows <- append(new_rows, list(new_row))
      }
    } else {
      # If no "/" is found, keep the row as is
      new_rows <- append(new_rows, list(df[i, , drop = FALSE]))
    }
  }
  # Combine the new rows into a data frame
  new_df <- do.call(rbind, new_rows)
  return(new_df)
}
```

## Run Split and Insert Row

```{r}
missing_df_71_12_13 <- split_and_insert_row(missing_df_71_12_13, "Title")
```


## Function No.2: 'remove_featuring_from_artist(df, "artist_col")'
- songlyrics.com doesn't play nice with artist features so we have to remove "featuring", "feat," and "ft."

```{r}
remove_featuring_from_artist <- function(df, artist_col) {
  df[[artist_col]] <- gsub("\\s*(feat|featuring|ft)\\..*", "", df[[artist_col]], ignore.case = TRUE)
  df[[artist_col]] <- trimws(df[[artist_col]])  # remove trailing whitespace
  return(df)
}
```

## Run Function No.2: removing feature

```{r}
missing_df_71_12_13 <- remove_featuring_from_artist(missing_df_71_12_13, artist_col = "Artist.s.")
View(missing_df_71_12_13)
```


## Function No.3: generate_song_lyrics()

```{r}
generate_songlyrics_urls <- function(df, artist_col, song_col, base_url_template) {
  # Format for URL: lowercase, strip special chars, replace spaces with hyphens
  format_for_url <- function(x) {
    x <- tolower(x)
    x <- gsub("\\s*(feat(uring)?|ft)\\.?\\s.*", "", x, ignore.case = TRUE)  # remove "featuring", "feat", "ft"
    x <- gsub("[^a-z0-9 ]", "", x)  # remove special characters
    x <- gsub(" +", "-", x)         # replace spaces with hyphens
    return(x)
  }
  
  # Extract base domain from URL template
  base_parts <- unlist(strsplit(base_url_template, "/"))
  base_domain <- paste(base_parts[1:3], collapse = "/")  # e.g. "https://www.songlyrics.com"
  
  # Generate URLs
  urls <- mapply(function(artist, song) {
    artist_url <- format_for_url(artist)
    song_url <- format_for_url(song)
    paste0(base_domain, "/", artist_url, "/", song_url, "-lyrics/")
  }, df[[artist_col]], df[[song_col]])
  
  # Return as data frame with column named 'lyrics_url'
  return(data.frame(lyrics_url = urls, stringsAsFactors = FALSE))
}
```


## Running the URL Generator Function

```{r}
url_for_missing_df_71_12_13 <- generate_songlyrics_urls(missing_df_71_12_13, "Artist.s.", "Title", "https://www.songlyrics.com")

```

## Combine with Missing_Dataframe

```{r}
replacement_vector <- url_for_missing_df_71_12_13$lyrics_url 

# Check if the number of NAs matches the length of the replacement vector
missing_df_71_12_13$lyrics_url[is.na(missing_df_71_12_13$lyrics_url)] <- replacement_vector

```

## Function to Scrape Lyrics

```{r}
scrape_lyrics_from_urls <- function(df, url_col, lyrics_col_name = "lyrics") {
  get_lyrics <- function(url) {
    tryCatch({ #tryCatch safely handle issues with Scraping returning NA in such cases like a Network Error
      webpage <- read_html(url) #Read HTML from URL
      
      # Extract lyrics from the page
      # Inspect the HTML structure of songlyrics.com to get the right CSS selector
      lyrics_node <- html_node(webpage, "#songLyricsDiv")
      
      # Get text and clean it up
      lyrics <- html_text(lyrics_node)
      lyrics <- trimws(lyrics)
      return(lyrics)
    }, error = function(e) {
      return(NA)  # Return NA on failure
    })
  }
  # Apply the scraping function to each URL
  df[[lyrics_col_name]] <- sapply(df[[url_col]], get_lyrics)

  return(df)
}
```

## Run Scrape Function

```{r}
missing_df_71_12_13 <- scrape_lyrics_from_urls(missing_df_71_12_13, "lyrics_url", lyrics_col_name = "lyrics")
```

## Merge 1971, 2012, 2013 to the main data frame
- formatting out duplicate data and existing NA valued columns from same years

```{r}
final_df_artist_lyrics <- merge(main_df_artist_lyrics, missing_df_71_12_13, by = intersect(names(main_df_artist_lyrics), names(missing_df_71_12_13)), all = TRUE)
final_df_artist_lyrics <- final_df_artist_lyrics[!is.na(final_df_artist_lyrics$Title), ]
final_df_artist_lyrics <- final_df_artist_lyrics[order(final_df_artist_lyrics$Year, final_df_artist_lyrics$Rank), ] 
final_df_artist_lyrics <- final_df_artist_lyrics[!is.na(final_df_artist_lyrics$Title), ]
```


## Taking Out Failed Lyric Scrape Songs from Main Data Frame

```{r}
missing_df_lyrics <- final_df_artist_lyrics %>% 
                     filter(is.na(final_df_artist_lyrics$lyrics))
```

## Deleting Old songlyrics URL 

```{r}
missing_df_lyrics <- remove_featuring_from_artist(missing_df_lyrics, "Artist.s.")
missing_df_lyrics$lyrics_url <- NA
```

## Remove Featuring from artist

```{r}
remove_featuring <- function(df, column) {
  df[[column]] <- sub("(?i)\\s*featuring.*$", "", df[[column]])
  return(df)
}
```

## Run Function for Featuring

```{r}
missing_format_df_lyrics <- remove_featuring(missing_df_lyrics, "Artist.s.")
print(missing_format_df_lyrics$Artist.s.)
```

## Regenerate Lyrics for this column

```{r}
missing_format_df_lyrics <- generate_songlyrics_urls(missing_df_lyrics, "Artist.s.", "Title", "https://www.songlyrics.com")

replacement_vector3 <- missing_format_df_lyrics$lyrics_url #replacement vector to be moved

missing_df_lyrics$lyrics_url[is.na(missing_df_lyrics$lyrics_url)] <- replacement_vector3  #combining data frames back into  line 214 df.
```


```{r}
scrape_azlyrics_from_df <- function(df, url_col, lyrics_col_name = "lyrics") {
  
  # Inner function to scrape a single URL
  scrape_lyrics <- function(url) {
    tryCatch({
      page <- read_html(GET(url, add_headers("User-Agent" = "Mozilla/5.0")))
      divs <- html_nodes(page, "div")
      main_col_index <- grep("class=\"col-xs-12 col-lg-8 text-center\"", as.character(divs))
      
      if (length(main_col_index) == 0) {
        message(sprintf("Could not locate main lyrics section for: %s", url))
        return(NA)
      }
      
      lyrics_index <- main_col_index + 6
      if (lyrics_index > length(divs)) {
        message(sprintf("Expected lyrics node not found at offset index for: %s", url))
        return(NA)
      }
      
      lyrics_text <- html_text(divs[[lyrics_index]])
      lyrics_text <- gsub("\r", "", lyrics_text)
      lyrics_text <- gsub("\n{2,}", "\n\n", lyrics_text)
      lyrics_text <- trimws(lyrics_text)
      return(lyrics_text)
    }, error = function(e) {
      message(sprintf("Scraping failed for: %s — %s", url, e$message))
      return(NA)
    })
  }
  
  # Apply the scraper to each URL in the specified column
  df[[lyrics_col_name]] <- sapply(df[[url_col]], scrape_lyrics)
  
  return(df)
}
```

```{r}
gen_missing_df_lyrics <- scrape_azlyrics_from_df(missing_df_lyrics, "lyrics_url", lyrics_col_name = "lyrics")
View(gen_missing_df_lyrics)
```

```{r}
write.csv(final_df_artist_lyrics, "ScrapedLyrics_AlmostFinish.csv", row.names = FALSE)
```

```{r}
closeAllConnections()
```
